\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra

### 2016-11-01 document/XSGRB_sample.tex
GRB120805A GRB120815A GRB120909A GRB121024A GRB121027A GRB161007A GRB161014A

\newcommand{\griz}{$g' r' i' z'$}
\newcommand{\JHK}{$JHK_{\rm{s}}$}
\newcommand{\gK}{$g' r' i' z' JHK_{\rm{s}}$}
\newcommand{\Msun}{$M_\odot$}
\newcommand{\lya}{Ly$\alpha$} 
\newcommand{\lyb}{Ly$\beta$} 
\newcommand{\hb}{H$\beta$} 
\newcommand{\ha}{H$\alpha$} 
\newcommand{\hg}{H$\gamma$} 
\newcommand{\hi}{\ion{H}{1}} 
\newcommand{\hei}{[\ion{He}{1}]} 
\newcommand{\oi}{[\ion{O}{1}]} 
\newcommand{\sii}{[\ion{S}{2}]} 
\newcommand{\siii}{[\ion{S}{3}]} 
\newcommand{\oii}{[\ion{O}{2}]$\lambda$3727} 
\newcommand{\oiii}{[\ion{O}{3}]$\lambda$5007}
\newcommand{\nii}{[\ion{N}{2}]} 
\newcommand{\neiii}{[\ion{Ne}{3}]} 
\newcommand{\NIii}{[\ion{Ni}{2}]} 
\newcommand{\feii}{\ion{Fe}{2}} 
\newcommand{\civ}{\ion{C}{4}} 
\newcommand{\cii}{\ion{C}{2}}
\newcommand{\mgi}{\ion{Mg}{1}} 
\newcommand{\mgii}{\ion{Mg}{2}} 
\newcommand{\ali}{\ion{Al}{1}} 
\newcommand{\alii}{\ion{Al}{2}} 
\newcommand{\aliii}{\ion{Al}{3}} 
\newcommand{\SIii}{\ion{Si}{2}} 
\newcommand{\SIiv}{\ion{Si}{4}} 
\newcommand{\znii}{\ion{Zn}{2}} 
\newcommand{\crii}{\ion{Cr}{2}} 
\newcommand{\mnii}{\ion{Mn}{2}} 
\newcommand{\tiii}{\ion{Ti}{2}} 
\newcommand{\caii}{\ion{Ca}{2}}

@@

@@

@@

	\subsection{GRB160804A (z = 0.736)}
	The data presented here also formed the basis of GCN \#
	19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}} and is not
	published elsewhere. Observations started 22.37 hours after the BAT trigger and
	lasted for 2.4ks. The afterglow continuum is detected across the entire
	spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii~and
	\alii~are found at $z = 0.736$. At the same redshift, emission lines from \oii,
	\oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch, lasting
	3.6ks, is obtained after the afterglow has faded, confirming the emission line
	detections.

>>>>

\newcommand{\griz}{$g' r' i' z'$}
\newcommand{\JHK}{$JHK_{\rm{s}}$}
\newcommand{\gK}{$g' r' i' z' JHK_{\rm{s}}$}
\newcommand{\Msun}{$M_\odot$}
\newcommand{\lya}{Ly$\alpha$} 
\newcommand{\lyb}{Ly$\beta$} 
\newcommand{\lyg}{Ly$\gamma$} 
\newcommand{\hb}{H$\beta$} 
\newcommand{\ha}{H$\alpha$} 
\newcommand{\hg}{H$\gamma$} 
\newcommand{\hi}{\ion{H}{1}} 
\newcommand{\hei}{[\ion{He}{1}]} 
\newcommand{\oi}{[\ion{O}{1}]} 
\newcommand{\sii}{[\ion{S}{2}]} 
\newcommand{\siii}{[\ion{S}{3}]} 
\newcommand{\oii}{[\ion{O}{2}]} 
\newcommand{\oiii}{[\ion{O}{3}]}
\newcommand{\nii}{[\ion{N}{2}]} 
\newcommand{\nv}{[\ion{N}{5}]} 
\newcommand{\neiii}{[\ion{Ne}{3}]} 
\newcommand{\NIii}{[\ion{Ni}{2}]} 
\newcommand{\feii}{\ion{Fe}{2}} 
\newcommand{\civ}{\ion{C}{4}} 
\newcommand{\cii}{\ion{C}{2}}
\newcommand{\mgi}{\ion{Mg}{1}} 
\newcommand{\mgii}{\ion{Mg}{2}} 
\newcommand{\ali}{\ion{Al}{1}} 
\newcommand{\alii}{\ion{Al}{2}} 
\newcommand{\aliii}{\ion{Al}{3}} 
\newcommand{\SIii}{\ion{Si}{2}} 
\newcommand{\SIiv}{\ion{Si}{4}} 
\newcommand{\znii}{\ion{Zn}{2}} 
\newcommand{\crii}{\ion{Cr}{2}} 
\newcommand{\mnii}{\ion{Mn}{2}} 
\newcommand{\tiii}{\ion{Ti}{2}} 
\newcommand{\caii}{\ion{Ca}{2}}

@@

	\subsection{GRB120805A (z $\sim$ 3.9 NEW!!!)}
	A separate reduction of this burst has been published in \citet{Kruhler2015},
	but not otherwise. Starting 9 days after the burst trigger, this is host
	observation and does not contain any afterglow continuum. In 3.6 ks integration
	time, we detect a faint continuum visible from 4500 \AA~and all the way through
	21000 \AA, in contrast to what is found previously. The continuum from 4500 -
	6000 \AA is detected at very low significance. If the drop at 4500 \AA~is the
	Lyman limit, this fits with Lyman alpha at $\sim$ 6000 \AA, giving $z \sim
	3.9$. The absence of nebular lines if due to \oii~falling in a telluric
	absorption band and the rest being shifted out of the wavelength coverage.

	\subsection{GRB120815A* (z = 2.358)} 
	Not a part of the statistical sample, thisburst also formed the basis of  GCN
	\# 13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
	published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
	trigger and consist of 2.4 ks integration. A bright afterglow continuum is
	detected across the entire spectral coverage of X-shooter, with a multitude of
	absorption lines superposed. Absorption features from the host at $z = 2.358$
	include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
	\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
	fine-structure lines from \NIii and \feii are exited local to the GRB.
	Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.

@@

	\subsection{GRB120909A (z = 3.929)}
	The data presented here has formed the basis of GCN \#
	13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
	published otherwise. A very rapid follow-up, starting only 1.7 hours after the
	BAT trigger, this 1.2 ks observation captures a very bright afterglow
	continuum, starting at 4500 \AA, signifying the onset of the Lyman limit for a 
	system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
	prominent absorption features in the form of \lya, \lyb, and \lyg, visible in
	the Lyman alpha forest. Metal absorption lines arising from \feii, \NIii,
	\SIii, \sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along
	with the corresponding fine-structure lines from (\feii$^*$, \SIii$^*$,
	\oi$^*$, \oi$^**$, \cii$^*$), securely anchoring the redshift of the host.

	\subsection{GRB121024A (z = 2.300)}
	The data presented here also formed the basis of GCN \#
	13890\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13890.gcn3}} and is published
	in \citet{Friis2015}. Also rapid, starting 1.8 hours after the \textit{Swift}
	trigger, a bright afterglow continuum is visible across all arms. A broad
	absorption feature from Lyman alpha, along with narrow lines from \civ, \SIii,
	\SIiv, \feii, \sii, and \alii, as well as fine-structure lines associated with
	\SIii$^*$ are all detected at $z = 2.300$, securely setting it as the redshift
	of the GRB.

	\subsection{GRB121027A (z = 1.773)}		
	The data presented here has formed the basis of GCN \#
	13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
	published otherwise. Starting 69.6 hours after the GRB trigger, that we detect
	the afterglow continuum a so high significance in all arms with 8.4 ks
	integration, testifies to the brightness of this burst. The concurrent
	identification of emission lines from \oiii~and absorption from \civ, \alii,
	\aliii, \mgi, \mgii, and \feii, tightly constrains the redshift of the burst to
	be $(z = 1.773$

@@

	\subsection{GRB160804A (z = 0.736)}
	The data presented here also formed the basis of GCN \#
	19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, but is not
	published elsewhere. Observations started 22.37 hours after the BAT trigger and
	lasted for 2.4ks. The afterglow continuum is detected across the entire
	spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii~and
	\alii~are found at $z = 0.736$. At the same redshift, emission lines from \oii,
	\oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch, lasting
	3.6ks, is obtained after the afterglow has faded, confirming the emission line
	detections.

	\subsection{GRB161007A (z =4.6 ??? NEW!!!)}
	This data has not been published elsewhere. Observations for GRB161007A started
	323 hours after the burst trigger and contains the potential host. 4 x 600 seconds
	of observations reveals a faint continuum rising abruptly above the noise at
	$\sim$ 6850 \AA~and continuing through 21000 \AA. A very low significance
	continuum is detected at shorter wavelengths, down to $\sim$ 6000 \AA. If the
	host is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha
	break and the absence of nebular emission lines is due to \oii being shifted
	out of the wavelength coverage. Alternatively, an early-type host at $z = 0.71$
	could exhibit the 4000 \AA break at 6000 \AA, but due to the preference of
	long-duration GRBs for star-forming galaxies, this is the least likely
	explanation, why we believe the high-z solution.

	\subsection{GRB161014A (z =2.823)}
	The data presented here also formed the basis of GCN \#
	20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}}, but is not
	published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8 ks of
	integration time captures the afterglow continuum across all three
	spectroscopic arms. A broad absorption trough due to Lyman alpha is visible,
	along with metal absorption features from \mgii, \SIii, \cii, \civ, \alii,
	\aliii, and	\feii, all at $z =2.823$. Similar to GRB140506 \citep{Fynbo2014}, a
	break in the continuum shape is detected bluewards of 6000\AA, possible
	signifying some anomalous form of extinction.
